HaploQA vs. R/qtl2 Haplotype Reconstructions

### 2022-12-22 draft_report.Rmd
updates 12/22

# directory to store input data retrieved from haploqa
data_dir <- file.path(root, config_sample$data_dir)
# directory to store input data for qtl2
qtl2_dir <- file.path(root, config_sample$qtl2_dir)
```

@@

1. Results for DO:

```{r message=FALSE}
do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

Below are the genotype visualizations for some individuals:

```{r}
plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(do_results[['ph_geno']], do_results[['ph_geno_haploqa']], do_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA')

# remove DO from memory, use gc()
```

@@

```{r message=FALSE}
cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

Genotype visualizations for some individuals:

```{r}
plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(cc_results[['ph_geno']], cc_results[['ph_geno_haploqa']], cc_results[['map']], ind = 50, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)

@@

Genotype visualizations for some individuals:

```{r}
plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(bxd_results[['ph_geno']], bxd_results[['ph_geno_haploqa']], bxd_results[['map']], ind = 5, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

```{r}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = T)
```

Genotype visualizations for some individuals:

```{r}
plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 1, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 1),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

```{r}
plot_onegeno_test(f2_results[['ph_geno']], f2_results[['ph_geno_haploqa']], f2_results[['map']], ind = 5, 
                          shift = TRUE, main = paste0('qtl2 - Geno-wide genotypes of individual ', 50),
                          sub = 'Left - qtl2, Right - HaploQA') 
```

@@

ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_gm <- ind_gm[!ind_gm %in% c('QU', 'SM', 'SX', 'T5', 'TG', 'TR', 'TS', 'TX', 'UF', 'UR', 'VA', 'VE', 'VG', 'VK', 'VN', 'W4', 'XS', 'Y8', 'Z8', 'ZG')] # screentime error
ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT', 'XZ', 'Y9')] # only one founder
ind_gm <- ind_gm[7:length(ind_gm)]
#gigamuga_results <- list()
#gm_pct_res <- list()
skipped_inds <- list()
for (ind in ind_gm) {
  print(ind)
  skip <- FALSE
  tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T), error = function(e) {skip <<- TRUE})
  if(skip) {
     skipped_inds <- c(skipped_inds, ind)
     next}
  #df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')

@@

```{r}
mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')

if (!file.exists(mini_res_dir)) {
  ### MiniMUGA pipeline
  minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
  url_domain <- 'http://haploqa-dev.jax.org/'
  marker_type <- 'MiniMUGA'
  html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
  url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
  summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
  ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
  ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
  ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
  ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
  minimuga_results <- list()
  mini_pct_res <- list()
  for (ind in ind_mini) {
    #ind <- 'JXN'
    print(ind)
    sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
    df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

```{r}
do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

@@

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA - percentage of marker difference between genail and haploqa') + xlab('model')

>>>>

# directory to store input data retrieved from haploqa
data_dir <- file.path(root, config_sample$data_dir)
# directory to store input data for qtl2
qtl2_dir <- file.path(root, config_sample$qtl2_dir)

print(config_sample)
```

@@

The pipeline to run haplotype reconstructions for an entire cross can be initiated using the function 'haplotype_reconstruction_pipeline'. Please set toggle 'samples_gen' = True if the genotyping data for individuals need to be (re)downloaded from HaploQA, and set to False otherwise. please set toggle 'qtl2_file_gen' = F if the input files for qtl2 read_cross need to be (re)generated, and False otherwise.

The function automatically saves the computed haplotype reconstruction results to an RDS file, which you can find in the /results/RDS/ directory of this repo. Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance and generate visualizations read the results from these RDS files.

1. Results for DO

```{r, eval=F}
do_results <- haplotype_reconstruction_pipeline('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

Below are the genotype visualizations for some individuals in DO:

a. Sample 6UY

![Karyotype plot - sample 6UY in DO](/Users/linc/Desktop/do_plot_1.png)

b. Sample 6WN

![Karyotype plot - sample 6WN in DO](/Users/linc/Desktop/do_plot_2.png)

@@

```{r, eval=F}
cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

a. Sample 35V

![Karyotype plot - sample 35V in CC](/Users/linc/Desktop/cc_plot_1.png)

b. Sample 6MV

![Karyotype plot - sample 6MV in CC](/Users/linc/Desktop/cc_plot_2.png)

@@

```{r, eval=F}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

a. Sample 5PD

![Karyotype plot - sample 5PD in F3](/Users/linc/Desktop/bxd_f3_plot1.png)

b. Sample 5R2

![Karyotype plot - sample 5R2 in F3](/Users/linc/Desktop/bxd_f3_plot2.png)

@@

```{r, eval=F}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

1. Sample 8RS

![Karyotype plot - sample 8RS in F2](/Users/linc/Desktop/f2_plot1.png)

b. Sample 8T3

![Karyotype plot - sample 8T3 in F3](/Users/linc/Desktop/f2_plot2.png)

@@

ind_gm <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_gm <- ind_gm[!ind_gm %in% c('RY', 'S8', 'SG', 'SR', 'Y2', 'YA', 'YJ', 'YT', 'XZ', 'Y9')] # only one founder

#gigamuga_results <- list()
#gm_pct_res <- list()
skipped_inds <- list()
for (ind in ind_gm) {
  print(ind)
  skip <- FALSE
  tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T), error = function(e) {skip <<- TRUE})
  if(skip) {
     skipped_inds <- c(skipped_inds, ind)
     next}
  #df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')

@@

#haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)

mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')

if (!file.exists(mini_res_dir)) {
  ### MiniMUGA pipeline
  minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
  url_domain <- 'http://haploqa-dev.jax.org/'
  marker_type <- 'MiniMUGA'
  html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
  url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
  summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
  ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
  ind_mini <- ind_mini[ind_mini!='JXX', 'JY6'] # screentime error, skip this one
  ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well

  minimuga_results <- list()
  mini_pct_res <- list()
  for (ind in ind_mini) {
    #ind <- 'JXN'
    print(ind)
    sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
    df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

```{r}
do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F)
f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

@@

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs qtl2 - percentage of marker discordance') + xlab('model')
